[Re] Chaos in a long-term experiment with a plankton community

### 2015-09-24 article/article.tex
added table 1 data

@@

The data are available as an Excel file supplement to
\href{http://onlinelibrary.wiley.com/doi/10.1111/j.1461-0248.2009.01391.x/abstract}{an
Ecology Letters publication} \textcite{Beninca2009}. The Excel file
contains several datasheets. Two are particularly important, as they are
the source of the raw data (one contains original species abundances,
the one with the nutrient concentrations). We saved these two datasheets
as comma separated value text files. In the code associated with this
reproduction, these data files are read from the associated github
repository.

Another datasheet in the Ecology Letters supplement contains transformed
variables, though this was not used in this reproduction, as we
reproduced the described transformations and applied them to the raw
data.

We also received some data direct from Steve Ellner, see below for
details.

@@

The R language and environment for statistical computing and graphics
was used to make the reproduction. Additional R packages required are:

\begin{verbatim}
library(tidyr)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(RCurl)
library(pracma)
library(oce)
library(tseriesChaos)
library(reshape2)
library(mgcv)
library(repmis)
\end{verbatim}

@@

Did our transformation give data matching the data we received from
Ellner, and that in the Ecology Letters supplement?

>>>>

\subsection{Scope of the reproduction}\label{scope-of-the-reproduction}

An attempt was made to reproduce the majority of the results in the
original article. Instances where we did not attempt to reproduce a
result are detailed below.

@@

The data are available as an Excel file supplement to
\href{http://onlinelibrary.wiley.com/doi/10.1111/j.1461-0248.2009.01391.x/abstract}{an
Ecology Letters publication} \textcite{Beninca2009}. The Excel file
contains several datasheets. Two were particularly important, as they
are the source of the raw data (one contains original species
abundances, the one with the nutrient concentrations). We saved these
two datasheets as comma separated value (csv) text files. In the code
associated with this reproduction, these data files are read from the
associated github repository.

Another datasheet in the Ecology Letters supplement contains transformed
variables (we also saved this as csv file, in order to use it in this
reproduction). We also received a dataset direct from Steve Ellner, see
below for details.

@@

The R language and environment for statistical computing and graphics
was used to make the reproduction. Additional R packages required are
specified in the code associated with this reproduction.

The code for this reproduction resides in an R markdown document, as
well as a source file containing some required functions. Some of the
code takes several minutes to run, so an intermediate data file is
provided with results from this code.

@@

The following transformation steps were used:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Time series shortened to remove long sequences of zeros.
\item
  Interpolation to create equally spaced observations in time series.
\item
  Fourth root transformation.
\item
  Detrending of five of the time series.
\item
  Rescaling to zero mean and unit standard deviation.
\end{enumerate}

The method for selecting the zeros to remove was unclear. In order to
reproduce this step, we removed the same data asin the original study,
by matching to the transformed data with zeros removed in the Ecology
Letters supplement Excel file mentioned above. All remaining
transformation steps were performed independently of this dataset. Our
reproduction of the transformed data closely matched the published
transformed data. \emph{This could result from us performing the
interpolation before the zero removal. Should correct this.}

The data received directly from Stephen Ellner was interpolated, but
without zeros removed. Our interpolated data, without zeros removed,
matched closely this data.

Correlations among species abundances presented in Table 1 of the
original article closely matched our reproduced correlations, calculated
from the transformed data with zeros removed.

\begin{longtable}[c]{@{}lcccccccccc@{}}
\caption{\label{tbl:table}Table caption }\tabularnewline
\toprule
& Bact & Harp & Ostr & N & P & Pico & Nano & Roti & Prot &
Cala\tabularnewline
\midrule
\endfirsthead
\toprule
& Bact & Harp & Ostr & N & P & Pico & Nano & Roti & Prot &
Cala\tabularnewline
\midrule
\endhead
Bact & & 0.03 & -0.24 *** & 0.05 & 0.18 *** & 0.03 & -0.17 *** & 0.3 ***
& -0.17 * & 0.22 ***\tabularnewline
Harp & & & 0.21 *** & -0.11 * & 0.03 & -0.09 & -0.06 & -0.05 & 0.14 &
0.01\tabularnewline
Ostr & & & & -0.16 ** & -0.06 & -0.06 & 0 & -0.05 & 0.18 * &
0.04\tabularnewline
N & & & & & 0.08 * & 0 & -0.02 & -0.04 & 0.03 & 0.04\tabularnewline
P & & & & & & -0.04 & 0.03 & 0.1 * & -0.02 & -0.08\tabularnewline
Pico & & & & & & & -0.17 *** & -0.03 & -0.21 * & 0.28 ***\tabularnewline
Nano & & & & & & & & -0.19 *** & 0.13 & -0.14 *\tabularnewline
Roti & & & & & & & & & -0.02 & -0.1 *\tabularnewline
Prot & & & & & & & & & & NA NA\tabularnewline
Cala & & & & & & & & & &\tabularnewline
\bottomrule
\end{longtable}
